---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data


```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data
1.Generalizability: since the survey is taken across all 50 states and other US territories, coordinated by the CDC with each state’s health agency, … - it should be enough to make random sample to make it generalizable to the US population.

Causality: since the BRFSS is an observational exercise, meaning there is no explicit random assignments to treatments, all relationships indicate a correlation but not a causation.
However, bias might exist due to the following reasons

Method of data collection is by phone. Hence, it is possible that there is underreporting on those that:

do not have a landline or mobile phone.

do not respond to phone surveys.

those who were not available to complete the survey (people who work/ preoccupied when the call was made )

Since there is no sure way to validate all answers, respondents could provide untruthful answers (or understate/overstate)

* * *

## Part 2: Research questions

**Research quesion 1:**
Is there a correlation between having a healthplan and general health? Any difference between gender?

**Research quesion 2:**
Is there is any correlation between how much time a person sleeps and diagnose of diseases such as heart disease or stroke

**Research quesion 3:**
Is there any relationship between income level and educational level? 


* * *

## Part 3: Exploratory data analysis

Is there a correlation between having a healthplan and general health? Any difference between gender?

genhlth: General Health

sex: Respondents Sex

hlthpln1: Have Any Health Care Coverage

Create a subset of data, omit all N/A

Create a table

Then plot the data.

**Research quesion 1:**

```{r}
qq1 <- select(brfss2013,genhlth,sex,hlthpln1) %>% na.omit()
dim(qq1)
```
```{r}
View(qq1)
prop.table(table(qq1$genhlth,qq1$hlthpln1),2)
```
```{r}
g <- ggplot(qq1) + aes(x=hlthpln1,fill=genhlth) + geom_bar(position = "fill")
g <- g + xlab("Have coverage") + ylab("Proportion") + scale_fill_discrete(name="Healthplan and general health")
g
```


```{r}
g <- ggplot(qq1) + aes(x=hlthpln1,fill=genhlth) + geom_bar(position = "fill") + facet_grid(.~sex)
g <- g + xlab("Gen health per gender") + ylab("Proportion") + scale_fill_discrete(name="Sectioned by gender")
g
```
Overall observation: people with health plan reports a lower ‘fair’ or ‘poor’ general health status.

People with no health plan reports more ‘good’ health status than those with health plan.

People with health plan reports a higher ‘excellent’ and ‘very good’ health status compare to those with no healthplan.

There is no noticeable discrepancy when sectioned into gender.
**Research quesion 2:**I wonder if there is any correlation between how much time a person sleeps and diagnose of diseases such as heart disease or stroke

cvdcrhd4: Ever Diagnosed With Angina Or Coronary Heart Disease

cvdstrk3: Ever Diagnosed With A Stroke

sleptim1: How Much Time Do You Sleep

Data have to be converted into numerical values.

```{r}
qq2 <- select(brfss2013,cvdcrhd4,cvdstrk3,sleptim1) 
qq2_pop <- select(qq2,cvdcrhd4,cvdstrk3,sleptim1) %>% na.omit() %>% filter(sleptim1 <= 16)
qq2_pop$cvdcrhd4 <- ifelse(qq2_pop$cvdcrhd4=="Yes", 1, 0)
qq2_pop$cvdstrk3 <- ifelse(qq2_pop$cvdstrk3=="Yes", 1, 0)
View(qq2_pop)
cor(qq2_pop)
```
plotting correlation

```{r}
library(corrplot)
```
```{r}
M <- cor(qq2_pop)
corrplot(M, method="ellipse")
```
From the results:

Sleep time and diagnosed with heart disease have a small negative correlation. It means that if one sleeps more, chances for heart disease is less.

Sleep time and diagnosed with a stroke has a small positive correlation. It means that if one sleeps more, chances for stroke can be higher.

**Research quesion 3:**Is there any relationship between income level and educational level?

income2: income level

educa: education level

sex: respondent’s sex
Create a subset, then filter out N/A values:

```{r}
qq3 <- select(brfss2013,educa,income2,sex)  %>%
filter(!is.na(income2), !is.na(sex), !is.na(educa))
```
Present totals of variables being analyzed
```{r}
qq3 %>% group_by(income2) %>%    summarise(count=n())
```
```{r}
qq3 %>% group_by(sex) %>%    summarise(count=n())
```
```{r}
qq3 %>% group_by(educa) %>%    summarise(count=n())
```

plot
```{r}
ggplot(data = qq3, aes(x = educa , y = income2 ))+
      geom_count () +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
    facet_grid(. ~  sex) +
      xlab("educa: Education Level") +
      ylab ("income2: Income Level")
      scale_x_discrete(labels = function(x) str_wrap(x, width = 5))
```
There seems to be positive correlation between education level and income level. Very little people from the never attend school bracket actually achive over 75,000 while a 30,000 people from each sex can achieve the 75,000 and above bracket if they attend 4 years college.

There is no noticeable discrepancy between the two sexes.

Ideally, I would love to have another bracket, naming people who attend further studies after undergraduate schools and more upper-level income bracket (perhaps above 100,000, above 1M)
